---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# 🗳 Plausi Package

**Predict votes and detect anomalies using R.**

<!-- badges: start -->
[![tic](https://github.com/machinelearningZH/plausi/workflows/tic/badge.svg)](https://github.com/machinelearningZH/plausi/actions)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

The Plausi package is designed for R-supported election forensics. It provides functions that enable the identification of statistical irregularities and anomalies in vote results.

Key features include:

- Robust outlier detection for small sample sizes and skewed distributions
- Calculation of differences between all possible combinations of turnout-levels (e.g., for systematic comparison of voter turnout across all voting districts)
- Prediction of expected results using machine learning algorithms (e.g., yes-vote proportions, voter turnout, etc.)

It serves as a basis for the ___PlausiApp___, which is used for vote result quality control in different cantons (TG / SG / ZH).

For the moment, the ___PlausiApp___ is made available upon request via a private Repo (mailto:wahlen@statistik.zh.ch). 

## Installation

You can install the development version of plausi from [GitHub](https://github.com/) with:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("machinelearningZH/plausi")
```


## Example

We will reference the pkgdown articles shortly.

```{r eval=FALSE}
library(plausi)
## basic example code
```


## Licensing
The package is licensed under the MIT license. 


## Project team
This is a joint project of the [Vote & Election-Team](https://www.zh.ch/de/direktion-der-justiz-und-des-innern/statistisches-amt/wahlen-und-abstimmungen.html) and [Team Data](https://www.zh.ch/de/direktion-der-justiz-und-des-innern/statistisches-amt/data.html) of the Statistical Office of the Canton of Zurich. Responsible: Simon Graf, Thomas Lo Russo and Thomas Knecht


## Feedback and contributing
We would love to hear from you. Please share your feedback and let us know how you use the code. You can [write an email](mailto:wahlen@statistik.zh.ch) or share your ideas by opening an issue or a pull requests.














